Interval Probability for Fuzzy Quantum Theories

### 2016-05-09 prob.tex
prob.tex Add comments from Gerardo...

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$
%% \yutsung{Use more physical or mathematical notation?}\\
of sequences of characters of length $N$ where each character is
either $X$ to indicate a needle crossing a line or $-$ to indicate a
needle not crossing a line. If $N=3$, the event that exactly 2 needles
cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{3}{8}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

@@

We begin with an analysis explaining how to combine intervals of
probability. Specifically if we have an event $E_1$ with an interval
of probability $[l_1,r_1]$ and another disjoint event $E_2$ with an
interval of probability $[l_2,r_2]$, what is the interval probability
of the event $E_1 \cup E_2$. The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_1+l_2,r_1+r_2]$ as
the combined probability, one can do find a much tighter interval if
information \emph{against} the event (i.e., information about the dual
event) is also taken into consideration.

@@

The mathematical framework above assumes that one has complete
knowledge of the events and their relationships. However, in many
practical situations, the structure of the event space is only
partially known and the precise dependence of two events on each other
cannot, a priori, be determined with certainty. In the quantum case,
this partial knowledge is compounded by the fact that there exist
non-commuting events which cannot happen simultaneously. To
accommodate these more complex situations, we abandon the sample
space~$\Omega$ and reason directly about events. A quantum probability
space therefore consists of just two components: a set of events
$\qevents$ and a probability measure
$\mu : \qevents \rightarrow [0,1]$. We give an example before giving
the formal definition.

\begin{example}[One-qubit quantum probability space] 
  Consider a one-qubit Hilbert space with states
  $\alpha \ket{0} + \beta \ket{1}$ such that
  $|\alpha|^2 + |\beta|^2 = 1$. The set of events associated with this
  Hilbert space consists of all projection operators. Each event is
  interpreted as a possible post-measurement state of a quantum system
  in current state $\ket{\phi}$. For example, the event $\proj{0}$
  indicates that the post-measurement state will be~$\ket{0}$; the
  event $\proj{1}$ indicates that the post-measurement state will
  be~$\ket{1}$; the event $\proj{\ps}$ where
  $\ket{\ps} = \frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$ indicates that the
  post-measurement state will be $\ket{\ps}$; the event
  $\mathbb{1} = \proj{0}+\proj{1}$ indicates that the post-measurement
  state will be a linear combination of $\ket{0}$ and $\ket{1}$; and
  the empty event $\mathbb{0}$ states that the post-measurement state
  will be the empty state. As in the classical case, a probability
  measure is a function that maps events to $[0,1]$: here is a partial
  specification of a possible probability measure:
\[\begin{array}{rcl}
\mu\left(\mathbb{0}\right) = 0, \quad
\mu\left(\mathbb{1}\right) =  1, \quad
\mu\left(\proj{0}\right) = 1, \quad
\mu\left(\proj{1}\right) = 0, \quad
\mu\left(\proj{\ps}\right) = 1/2, \quad \ldots
\end{array}\]
Note that, similarly to the classical case, the probability of
$\mathbb{1}$ is 1 and the probability of collections of orthogonal
events (e.g., $\proj{0}+\proj{1}$) is the sum of the individual
probabilities. In contrast, a collection of non-orthogonal events
(e.g., $\proj{0}$ and $\proj{\ps}$) is not itself an event.
In the classical example, we argued that each probability measure is
uniquely determined by two actual coins. A similar (but much more
subtle) argument is valid also in the quantum case. By postulates of
quantum mechanics and Gleason's theorem, it turns out that for large
enough quantum systems, each probability measure is uniquely
determined by an actual quantum state.
\qed\end{example}

>>>>

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles one needle at a time, and observe the number
$X$ of needles that cross a line. The sample space can be expressed as the set
$\{X,-\}^N$
%% \yutsung{Use more physical or mathematical notation?}\\
of sequences of characters of length $N$ where each character is
either $X$ to indicate a needle crossing a line or $-$ to indicate a
needle not crossing a line. If $N=3$, the event that exactly 2 needles
cross lines~$\{-XX,X-X,XX-\}$ has
probability~$\frac{3}{8}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

@@

We begin with an analysis explaining how to combine intervals of
probability. Specifically if we have an event $E_1$ with an interval
of probability $[l_1,r_1]$ and another disjoint event $E_2$ with an
interval of probability $[l_2,r_2]$, what is the interval probability
of the event $E_1 \cup E_2$. The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_1+l_2,r_1+r_2]$ as
the combined probability, one can do find a much tighter interval if
information \emph{against} the event (i.e., information about the complement
event) is also taken into consideration.

@@

The mathematical framework above assumes that one has complete
knowledge of the events and their relationships, and moreover, there exists a
predetermined set of properties.
However, in many practical situations, the structure of the event space is only
partially known and the precise dependence of two events on each other
cannot, a priori, be determined with certainty. In the quantum framework,
this partial knowledge is compounded by the fact that there exist
non-commuting events which cannot happen simultaneously. To
accommodate these more complex situations, we abandon the sample
space~$\Omega$ and reason directly about events. A quantum probability
space therefore consists of just two components: a set of events
$\qevents$ and a probability measure
$\mu : \qevents \rightarrow [0,1]$. We give an example before giving
the formal definition.

\begin{example}[One-qubit quantum probability space] 
  Consider a one-qubit Hilbert space with states
  $\alpha \ket{0} + \beta \ket{1}$ such that
  $|\alpha|^2 + |\beta|^2 = 1$, $\alpha ,\beta \in \C$. The set of events
  associated with this Hilbert space consists of all projection operators. Each event is
  interpreted as a possible post-measurement state of a quantum system
  in current state $\ket{\phi}$. For example, the event $\proj{0}$
  indicates that the post-measurement state will be~$\ket{0}$; the
  event $\proj{1}$ indicates that the post-measurement state will
  be~$\ket{1}$; the event $\proj{\ps}$ where
  $\ket{\ps} = \frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$ indicates that the
  post-measurement state will be $\ket{\ps}$; the event
  $\mathbb{1} = \proj{0}+\proj{1}$ indicates that the post-measurement
  state will be a linear combination of $\ket{0}$ and $\ket{1}$; and
  the empty event $\mathbb{0}$ states that the post-measurement state
  will be the empty state. As in the classical case, a probability
  measure is a function that maps events to $[0,1]$: here is a partial
  specification of a possible probability measure:
\[\begin{array}{rcl}
\mu\left(\mathbb{0}\right) = 0, \quad
\mu\left(\mathbb{1}\right) =  1, \quad
\mu\left(\proj{0}\right) = 1, \quad
\mu\left(\proj{1}\right) = 0, \quad
\mu\left(\proj{\ps}\right) = 1/2, \quad \ldots
\end{array}\]
Note that, similarly to the classical case, the probability of
$\mathbb{1}$ is 1 and the probability of collections of orthogonal
events (e.g., $\proj{0}+\proj{1}$) is the sum of the individual
probabilities. In contrast, a collection of non-orthogonal events
(e.g., $\proj{0}$ and $\proj{\ps}$) is not itself an event.
In the classical example, we argued that each probability measure is
uniquely determined by two actual coins. A similar (but much more
subtle) argument is valid also in the quantum case. By postulates of
quantum mechanics and Gleason's theorem, it turns out that for large
enough quantum systems, each probability measure is uniquely
determined by an actual quantum state.
\qed\end{example}
